importing the required package
```{r}
require(tidyverse)

mydata <- read.csv("C:/Users/himanshu/Desktop/R analysis/mlb1.csv")
```

```{r}
temp <- rowSums(mydata[c("frstbase", "scndbase", "shrtstop", "thrdbase", "outfield", "catcher")])
min(temp)
max(temp)
```
As the result is 1 for both min and max values, we have confirmed that every player has been assigned exactly one position.

lets create two variables "position" and "race" which summerizes the respective information in one variable.

```{r}
mydata$position <- "First Base"
mydata$position[mydata$scndbase == 1] <- "Second Base"
mydata$position[mydata$shrtstop == 1] <- "Short Stop"
mydata$position[mydata$thrdbase == 1] <- "Third Base"
mydata$position[mydata$outfield == 1] <- "Outfield"
mydata$position[mydata$catcher == 1] <- "Catcher"
# ensure it is a factor variable
mydata$position <- as.factor(mydata$position)


mydata$race <- "White"
mydata$race[mydata$hispan == 1] <- "Hispanic"
mydata$race[mydata$black == 1] <- "Black"
mydata$race <- as.factor(mydata$race)

summary(mydata$position)
summary(mydata$race)
summary(mydata$salary)
```

we can get mean, median, sd, max, min, variance by directly using the funtions.

# First Pipe
```{r}
mydata %>% group_by(position) %>% summarise(mean(salary))
```
%>% is the piping operator

```{r}
mydata %>% group_by(position) %>% summarise(number = length(salary), avg.salary = mean(salary))
```


## filter()

```{r}
mydata %>% filter(yrsallst == 0) %>% group_by(position) %>% summarise(number = length(salary), avg.salary = mean(salary))
```


## arrange()

```{r}
mydata %>% filter(yrsallst == 0) %>% group_by(position) %>% summarise(number = length(salary), avg.salary = mean(salary)) %>% arrange(avg.salary)

```

## double pivot tables

```{r}
mydata %>% group_by(position, race) %>% summarise(avg.salary = mean(salary))
```


## spread() and arrange()

```{r}
mydata %>% group_by(position, race) %>% summarise(avg.salary = mean(salary)) %>% spread(race, avg.salary)
```

```{r}
mydata$AS <- (mydata$yrsallst>0)

mydata %>% group_by(AS, position, race) %>% summarise(avg.salary = mean(salary)) %>% spread(race, avg.salary) %>% arrange(AS)
```

